\documentclass[12pt]{article}

\title{Hello Agda}
\author{Alexander Ben Nasrallah}

\usepackage{neosimsim}
\usepackage{neosimsim-a4paper}

\usepackage[backend=biber]{biblatex}
\addbibresource{bibliography.bib}

\usepackage{titlesec}
\newcommand{\sectionbreak}{\clearpage}

\usepackage{fancyvrb}
\newcommand{\ignore}[1]{}

\newtheorem{exercise}{Exercise}

\DefineVerbatimEnvironment%
  {code}{Verbatim}
  {} % Add fancy options here if you like.

\DefineVerbatimEnvironment%
  {neolisting}{Verbatim}
  {frame=single} % Add fancy options here if you like.

\DefineVerbatimEnvironment%
  {haskell}{Verbatim}
  {frame=single,label=\textrm{Haskell}}

% https://tex.stackexchange.com/questions/18322/using-fancyvrb-verbatim-environment-from-within-newenvironment
\usepackage{fancyvrb,caption,floatrow}
\DeclareNewFloatType{neolistingfloat}{placement=t,fileext=neolisting,name=Listing}
\captionsetup[neolistingfloat]{}

\newenvironment{neocaplisting}[2]
{\VerbatimEnvironment%
 \captionof{neolistingfloat}{#2}\ifx\relax#1\relax\else\label{#1}\fi%
\begin{Verbatim}[frame=single]}
{\end{Verbatim}}

\begin{document}
\maketitle{}
\begin{abstract}
  These are notes on a workshop to introduce Agda, for the Berlin Haskell User Group.
  This workshop should be done interactivly and possible limited to 2 hours
  per session. I dont want to enforce the feeling of a real lecture.
  This workshop is most of all meant to make fun.
\end{abstract}

\tableofcontents{}\newpage{}

A few question to adjust the focus of the workshop.
\begin{itemize}
  \item Who has used Haskell before?
  \item Who has used Agda before?
  \item Who has used a dependent typed programming language before?
  \item Who used a proof system before?
\end{itemize}

\input{installation.md}
\input{hello-world.md}
\input{src/Naturals.lagda}
\input{Vector.lagda}

\section{Coming From Haskell}
\subsection{Unicode and Mixfix}
\input{unicode.lagda.tex}

\printbibliography{}
\end{document}
